### 2019-10-14 ch8_discussion/content.tex
update discussion

We developed \texttt{dyngen}, a simulator of single cells to benchmark and stress-test computational tools for single-cell omics (Chapter~\ref{chap:dyngen}).
It supports a selection of existing experimental protocols, but can also simulate experiments for permitted by current technologies, such as generating omics profiles from the same cell at multiple timepoints.
This data can help kick-start emerging domains with low data availability more safely by allowing method developers to test their method before numerous datasets become publicly available.

\texttt{dyngen} has already been used successfully to evaluate trajectory inference\cite{saelens_comparisonsinglecelltrajectory_2019}, trajectory alignment\cite{vandenberge_trajectorybaseddifferentialexpression_2019}, and network inference\cite{pratapa_benchmarkingalgorithmsgene_2019} methods. Furthermore, we showed that it can also be used to evaluate differential network inference methods, and trajectory alignment methods.

@@

For our second goal, it seems too early to tell whether our research had any effect on self-assessments of TI method developers. We did made our pipeline, datasets, metrics, and containerised wrappers of TI methods publicly available for developers to use.
However, so far we do not observe an increase in developers performing quantitative benchmarks. We hypothesise causal reasons for this phenomenon and provide solutions in order to spur TI developers to perform more self-assessments (Chapter~\ref{chap:selfassessment}).

\section{A toolkit to infer, visualise and interpret single-cell trajectories}

The previous two works necessitated developing a tool for visualising trajectory data. We extended this implementation into a full toolkit, named \texttt{dyno}, for inferring and analysing trajectories (Chapter~\ref{chap:dyno}). The toolkit allows performing downstream analyses such detecting differentially expressed genes along transitions in the trajectory, annotating the trajectory, and comparing multiple trajectories in a common dimensionality reduction. A major benefit of \texttt{dyno} is that it allows utilising any of the 50 TI methods we wrapped as part of the comparison, without having to install dependencies for every TI method separately.

@@

@@

Being able to study how cells progress over time is an analysis as fundamental as being able to cluster cells or identify differentially expressed genes between two groups. We showed that for analysing datasets containing linear trajectories, \texttt{SCORPIUS} outperforms all other TI methods tested as part of our TI benchmark.

@@

While \texttt{incgraph} solved a very specific and interesting problem, and was eventually was published as a peer-reviewed article, it has not succeeded in making a direct impact on scientific research as of yet. Perhaps, at some point, a researcher will come across \texttt{incgraph} and find that it exactly solves his or her problem.

@@

\textbf{Git}\cite{torvalds_gitfastversion_2005} is a code-revision system that allows multiple users to collaborate on developing code and keep track of what changes were made by whom. 
Since Wouter and I were often working closely together on a specific part of our software, Git saved us a lot of time by merging together changes developed in parallel. Only in few cases did we need to intervene manually to merge our changes.
Additionally, Github.com allowed us not only to collaborate with each-other, but also correspond and collaborate with other software developers from many different research groups, using Github Issues to discuss problems with other researchers, and Github Pull Requests to contribute code to open-source projects.

Together, we published over 20'000 code contributions (commits) across 50+ software packages\cite{cannoodt_developmentdynverse_2019}. Since many of these packages depend on one another, it is inevitable that changes made to one package would break another package. We wrote many, many unit tests (and still too few), and let these tests automatically be executed on \textbf{Travis CI}\cite{traviscigmbh_traviscitest_2011}. This way, when we pushed breaking changes to Github, Travis CI would notify that our commit has resulted in one or more unit tests failing. While sometimes it is a hassle to set up, Travis CI has prevented us countless times from generating faulty results due to a faulty underlying function.

>>>>

We developed \texttt{dyngen}, a simulator of single cells to benchmark and stress-test computational tools for single-cell omics (Chapter~\ref{chap:dyngen}).
The \texttt{dyngen} software supports a selection of existing experimental protocols, and can also simulate experiments that are possible with the current technologies, such as generating omics profiles from the same cell at multiple timepoints.
The generated data can help kick-start emerging domains with low data availability more safely by allowing software developers to test their method before numerous datasets become publicly available.

\texttt{dyngen} has already been successfully used to evaluate trajectory inference\cite{saelens_comparisonsinglecelltrajectory_2019}, trajectory alignment\cite{vandenberge_trajectorybaseddifferentialexpression_2019}, and network inference\cite{pratapa_benchmarkingalgorithmsgene_2019} methods. Furthermore, we showed that it can also be used to evaluate differential network inference methods, and trajectory alignment methods.

% Eventueel kort ook beschrijven waar je denkt dat het nog van toepassing kan zijn? The use if dyngene is not limited to.... but might also,....
% Eventueel ook: waar zie je nog verbeter/uitbreidingspunten in je methode. 
% Wat zijn de huidige limitaties. Bv. Is er een tijdgebonden limitatie. bv. Stel je wilt data van 10.000 cellen genereren maar bent een maand bezig, etc.

@@

For our second goal, it seems too early to tell whether our research had any effect on self-assessments of TI method developers. We did made our pipeline, datasets, metrics, and containerised wrappers of TI methods publicly available for developers to use.
However, so far we do not observe an increase in developers performing quantitative benchmarks. We hypothesise causal reasons for this phenomenon and provide solutions in order to spur TI developers to perform more self-assessments (Chapter~\ref{chap:selfassessment}).
% "We hypothesise." Where/what do you hypothise? In your paper? Dan zou ik verwijzen naar deze sectie.
% Eventueel ook: Is there room for improvement for existing methods (deze zullen vermoedelijk onder een andere naam gepubliceerd dienen te worden)
% Dienen eerdere publicaties in vraag te worden gesteld (indien minder goede methods gebruikt werden?
% Kan dit uitgebreid worden naar andere methoden?
% Is dyngen niet ontstaan omdat je nood had aan betere datasets? Zou je in dit opzicht niet dit deeltje eerst plaatsen en verwijzen naar dyngen?
% Misschien ook vermelden dat de publicatie een aanzienlijke support verwacht van jullie team na publicatie. Dat de hoge nood gepaard gaat met een grote vraag. 
\section{A toolkit to infer, visualise and interpret single-cell trajectories}

The previous two works necessitated developing a tool for visualising trajectory data. We extended this implementation into a full toolkit, named \texttt{dyno}, for inferring and analysing trajectories (Chapter~\ref{chap:dyno}). The toolkit allows performing downstream analyses such as detecting differentially expressed genes along transitions in the trajectory, annotating the trajectory, and comparing multiple trajectories in a common dimensionality reduction. A major benefit of \texttt{dyno} is that it allows utilising any of the 50 TI methods we wrapped as part of the comparison, without having to install dependencies for every TI method separately.

@@

% Are we talking about R-packages?

@@

Being able to study how cells progress over time is in terms of analysis as fundamental as being able to cluster cells or to identify differentially expressed genes between two groups. We showed that for analysing datasets containing linear trajectories, \texttt{SCORPIUS} outperforms all other TI methods tested as part of our TI benchmark.

@@

While \texttt{incgraph} solved a very specific and interesting problem, and was eventually published as a peer-reviewed article, it has not succeeded in making a direct impact on scientific research as of yet. Perhaps, at some point, a researcher will come across \texttt{incgraph} and find that it exactly solves his or her problem.

% Of je dient voor te stellen bij welke dataset het goed van toepassing is en het te promotor onder de Vlaamse research gemeenschap?

@@

\textbf{Git}\cite{torvalds_gitfastversion_2005} is a code-revision system that allows multiple users to collaborate on developing code and keep track of what changes were made by whom. 
Since Wouter and I were often working closely together on a specific part of our software, Git saved us a lot of time by merging together changes developed in parallel. Only in few cases did we need to intervene manually to merge our changes.
Additionally, Github.com allowed us not only to collaborate with each-other, but also correspond and collaborate with other software developers from many different research groups, using Github Issues to discuss problems with other researchers, and Github Pull Requests to contribute code to open-source projects. 
% Misschien ook Wouter zijn achternaam gebruiken. Klinkt wat professioneler :-p
Together, we published over 20'000 code contributions (commits) across 50+ software packages\cite{cannoodt_developmentdynverse_2019}. Since many of these packages depend on one another, it is inevitable that changes made to one package would break another package. We wrote many, many unit tests (and still too few), and we let these tests automatically be executed on \textbf{Travis CI}\cite{traviscigmbh_traviscitest_2011}. This way, when we pushed breaking changes to Github, Travis CI would notify that our commit has resulted in one or more unit tests failing. While sometimes it is a hassle to set up, Travis CI has prevented us countless times from generating faulty results due to a faulty underlying function.
